### 2017-01-27 paper1.tex
Corrections made to Introduction
Paper1 fixed to be published version

%\titlerunninghead{Solar Control of Polar Irregularities}

%\authoraddr{L. J. Lamarche and R. A. Makarevich, Geophysical Institute, University of Alaska Fairbanks, 903 Koyukuk Drive, PO Box 757320, Fairbanks, AK, 99775-7320, USA.
%(rmakarevich@alaska.edu)}

%\linenumbers*[1]
%\begin{document}

@@

%\authors{Leslie J. Lamarche\altaffilmark{1} and Roman A. Makarevich \altaffilmark{1}}
%\vspace{-1.5cm}

%\altaffiltext{1}{Geophysical Institute and Department of Physics, University of Alaska Fairbanks, Fairbanks, AK, USA.}

@@

%\begin{article}
\section{Introduction}
\label{Sec:Intro}

@@

Observations of decameter-scale irregularities with SuperDARN have an advantage of significant datasets accumulated under a wide variety of observational conditions. SuperDARN utilizes radio refraction in the High Frequency (HF) band to observe \(F\) region echoes or backscatter from decameter-scale irregularities when radar beam achieves orthogonality with local magnetic field (orthogonality condition). Because of importance of propagation conditions for HF radar observations, the echo occurrence rates observed by SuperDARN are taken as a lower limit of the irregularity occurrence rates \citep{Chisham2007}.

Several recent SuperDARN studies have found clear seasonal and diurnal trends in the \(F\) region echo occurrence that formed a distinct pattern which appeared to be qualitatively similar to that of the solar zenith angle (SZA) variations \citep{Kane2012,Ghezelbash2014a,Ghezelbash2014b}. However, it has been challenging to determine how occurrence explicitly depends on SZA for all seasonal and time sectors. The nighttime occurrence taken separately was demonstrated to be closely related to SZA-controlled \(F\) region density  variations \citep{Kane2012}, but no similarly-clear relationship has been demonstrated for datasets including daytime points.

One difficulty is the need to consider propagation effects of plasma density. These arise because of \(F\)-region density controlling the amount of refraction and either under- or over-refraction reducing echo occurrence when orthogonality condition is not met \citep{Milan1997,Danskin2002}. This is also related to how echo occurrence is generally considered a lower limit on irregularities (described above). There is some ambiguity about whether changing irregularity production or detection by the radar is most responsible for any observed trends.

The production/detection ambiguity issue can be addressed by considering how both occurrence and distance from the radar (slant range) change in responses to various factors \citep{Kane2012,Ghezelbash2014b}. This is because range variations are expected to be almost entirely due to propagation effects and the differences in responses are therefore mostly likely due to changing production. In order to claim production as the dominant source of occurrence variability, occurrence analysis must explicitly considers the echo location. In the current study this approach is employed to develop a comprehensive model of occurrence location, quantify echo location trends, and explicitly consider these trends when developing occurrence estimates. This is in contrast with previous studies that considered echoes detected in the entire (or fixed part) of the radar's field-of-view (FoV) \citep[e.g.][]{Kumar2011,Kane2012}.

@@

The primary instrument used in this study is the SuperDARN McMurdo radar located at McMurdo Station, Antarctica (77.88\(\deg\) S, 166.73\(\deg\) E, geographic) \citep{Bristow2011}. Using a standard SuperDARN 3-letter radar code, this radar is referred to as MCM. SuperDARN is an array of coherent HF radars used to study processes in the Earth's ionosphere and magnetosphere. Radars are located at mid and high latitudes in both the Northern and Southern Hemispheres. Radars measure a 17-lag auto-correlation function (ACF) from which estimates of the Doppler velocity, power, and spectral width of ionospheric echoes in 75--100 range bins for each of the 16--24 radar beams are obtained. Each radar beam spans \(\sim3.25\deg\) horizontally and \(\sim30\deg\) vertically, with the entire FoV of \(\sim52\deg\)--\(78\deg\) in azimuth. Range bins typically start at 180 km and have a length of 15 km or 45 km and a radar scans through all beams in 60 s or 120 s depending on the mode of operation. Other details on operations and technical characteristics of SuperDARN radars are given by \citet{Chisham2007}.

The magnetic latitude (MLAT) of MCM is 80.0\(\deg\) S, based on altitude-adjusted corrected geomagnetic (AACGM) coordinate system with 2010 coefficients \citep[see recent paper by][]{Shepherd2014}. The MCM FoV covers the southern polar cap and beam 14 of MCM is most conjugate with beam 1 of the SuperDARN Rankin Inlet (RKN) radar that overlooks the northern polar cap. The current study focuses on MCM observations, but in order to facilitate future comparisons with conjugate observations, the data from beam 14 of MCM are considered. MCM has been in routine operations since February 2010 and the current study has considered the data from the first 4 years of operations 2010--2013.

Standard SuperDARN criteria were used to eliminate ground scatter (\(\left| V-\Delta V \right| <30\) m/s and \(\left| W-\Delta W \right| <35\) m/s), where \(V\) is the Doppler velocity, \(W\) is the spectral width, and \(\Delta V\) and \(\Delta W\) represent uncertainties in these quantities. In addition, measurements with SNR \(<\) 3 dB or \(W>\) 500 m/s were eliminated to reduce contribution from interference and other spurious echoes. In addition to standard operation modes, each SuperDARN radar has a set time periods where the radar is run in special modes. In order to have as continuous of a dataset as possible for the statistical analysis, data were used both from the standard modes and from special modes when possible. Modes where the beam 14 was over-sampled (beam camping modes) or not sampled at all were not used.

In this study, we also employed the IMF measurements taken from the high-resolution OMNI database using spacecraft-specific dataset for the WIND satellite. This was the OMNI dataset that had the largest number of valid measurements throughout the period of interest 2010--2013. All data in the OMNI database are time shifted from the spacecraft location to Earth's bow shock nose (BSN). The data were further shifted by 9--12 min to account for propagation from the BSN to the high-latitude ionosphere using a method described by \citet{Khan1999} that was also employed by \citet{Makarevich2012}. The IMF By and Bz components were averaged over 10-min intervals. Any periods with IMF information missing were excluded from the analysis.

The model electron densities were also used in the current study. The electron densities were found from the International Reference Ionosphere (IRI) 2007 model \citep{Bilitza2008}. The IRI model is an empirical model of the ionosphere based on time of day, date, geographic location, altitude, and solar conditions, with electron density near the \(E\) and \(F\) region peaks being of particular interest for this study.  Using ionosonde terminology, these are also referred to below as model NmE and NmF2.

@@

In order to examine diurnal, seasonal, and solar cycle trends in echo occurrence, the monthly average occurrence was found for each universal time (UT) 2-min interval and for each 45-km range bin. If measurements were conducted at a 2-min resolution, each radar cell had an occurrence of 1 or 0. For a higher resolution, the values were weighted accordingly so that if an echo was observed for every possible time within the 2-min interval, the overall echo occurrence was still 1. The data for each time/range cell were averaged over the entire month. Figures \ref{fig:month_avg_occ}a--d show monthly average occurrence expressed in percentage for four selected months in year 2010 versus UT and slant range. The four months presented were chosen to be representative of different seasons. Figures \ref{fig:month_avg_occ}e--h on the right similarly shows the monthly average peak \(F\)-region electron density NmF2 obtained from the IRI model. The cosine white contours in both left and right panels show the approximate range position of the \(F\)-region echo band for each month determined as described below.

\begin{figure}
\includegraphics[width=12.5 cm,angle=90]{p1f1.pdf}
\caption{Monthly average plots of echo occurrence (left) and peak electron density (right) for 2010.  Four months are selected (one for each season) to show seasonal variations. In Figures \ref{fig:month_avg_occ}a--\ref{fig:month_avg_occ}d,  the pink line shows the 25\% occurrence contour that is used to define the echo band region. Black dots are the midpoints between the upper and lower portion of this contour.  A cosine function (white line) is fit to these points and used to model the echo band. The dashed white lines are the upper and lower bounds of the model band. The top right corner of each panel shows the parameters that characterize the band for that month: \(A\) is amplitude of cosine function (in km), \(B\) is time shift (in hrs), \(C\) is range shift from 0 (in km), and BW is the band width (constant distance between the upper and lower dashed white lines).  In Figures \ref{fig:month_avg_occ}e--\ref{fig:month_avg_occ}h the same echo location model is also given (white lines) along with expected echo locations based on raytracing simulations (dark red line).}
\label{fig:month_avg_occ}
\end{figure}

Figures \ref{fig:month_avg_occ}a--d show a distinct \(F\)-region band of high occurrence in every month at ranges 500--1500 km. Some \(E\)-region echoes are also observed at shorter ranges 180--300 km in June at 01--06 UT and in December at 17--24 UT. The \(F\)-region band range position and peak occurrence within the band was changing most clearly throughout the day in March and September, i.e. diurnal variation in both parameters is observed clearly in spring and fall months. To highlight the high echo occurrence band, a contour of 25\% occurrence is plotted in each left panel by the pink line and this shows a clear diurnal, cosine-like variation with UT in March and September.

The midpoints were also found between the top and bottom parts of the pink 25\% occurrence contour and the cosine function of the form
\begin{equation}\label{E:Model}
R\left(t\right) = A \cos\left[2\pi\left(t-B\right)/24\right] + C
\end{equation}
was fitted to these points. Here \(R\) is the range of the center of the band in kilometers, \(t\) is the time in hours, and  \(A\), \(B\), and \(C\) are fit coefficients with units of kilometers, hours, and kilometers, respectively. The white solid line in each panel of Figure \ref{fig:month_avg_occ} shows the resulting fit \(R\left(t\right)\). The white dashed lines represent the upper and lower limits of the band.  They were calculated by first finding the average distance \(\sigma\) between points in the 25\% occurrence contour and the fitted cosine at the same time of day. The original cosine fit was then shifted up or down by this amount as \(R\left(t\right) \pm \sigma\) to give the upper and lower limits of the band.  These limits match the pink 25\% occurrence contour well because all the points in the pink contour are approximately the same distance away from the central fit (white solid line). The band width (BW) could then be defined as \(2\sigma\). The fit parameters and band width are given in the top-right corner of Figures \ref{fig:month_avg_occ}a--d. For the rest of this study, ``echoes within the band'' refer to data points that fall between the upper and lower white dashed lines. From Figures \ref{fig:month_avg_occ}a--d, both the center position and upper/lower bounds for the band shown by the white lines capture the band and its diurnal variation in range very well for all months. Therefore, these estimates are used below to quantify the echo band behavior, particularly in context of expected electron densities.

MCM is oriented in such a way that it samples daytime (nighttime) ionosphere in roughly the first (second) half of the day in UT. The electron density on the right of Figure \ref{fig:month_avg_occ} clearly shows higher daytime densities that reach a peak near 05 UT at farther ranges and lower nighttime densities that reach a minimum near 17 UT. The \(F\)-region echo band thus was closer to the radar in the daytime (high densities) and farther away at night (low densities). This feature is most clear for the March and September datasets. It is much less clear in June, although one can still see a portion of the cosine-like variation at 02--10 UT in June in the pink contour. In this month, the white contours are mostly extrapolation outside of this interval, but they fit into the overall diurnal/seasonal pattern well (see also next analysis).  In December, the band is flat and diurnal variation is not evident.

In terms of diurnal variation of echo occurrence, there was a minimum occurrence at nighttime and a maximum in daytime hours. The minimum was much deeper and more extended in June so that essentially no echoes were observed at 12--24 UT. At this time, the expected location of echoes based on the cosine-fit model (white lines) is farther away from MCM than in any month and any UT interval, but no echoes are observed either at these ranges or any other ranges. This is in contrast with March and September, where occurrence is reduced at nighttime but not to zero. Overall, Figure \ref{fig:month_avg_occ} shows that range position of the band is largely controlled by the density conditions, while the echo occurrence variation is perhaps more complicated.

The final observation about Figure \ref{fig:month_avg_occ} concerns the results of raytracing simulations that were conducted for each hour of day 15 in each of the four months shown in Figure \ref{fig:month_avg_occ}. In these simulations, IRI densities in a 2D simulation domain aligned with beam 14 of MCM were used. The range locations where orthogonality condition was expected are plotted in Figures \ref{fig:month_avg_occ}e--h by the dark red line. The gaps during the nighttime in these lines in all months but December indicate that no orthogonality was expected at any range in MCM beam 14 during these periods. The red line is located between the solid and bottom dashed white lines, indicating a reasonable agreement between the expected and observed backscatter locations. The important exception is seen, however, for nighttime periods (or, more accurately, whenever densities were lower than \(\sim1.5\times10^5\) cm\(^{-3}\), dark blue colors). During those periods, no orthogonality was expected based on IRI densities, but significant (albeit reduced) amount of backscatter was observed in March and September. This implies a significant source of nighttime plasma density that is unaccounted for by the IRI model.

The same analysis was performed for the entire period of interest 2010--2013 and the band characteristics are presented in Figure \ref{fig:year_line}. It shows the band (a) position amplitude \(A\), (b) time shift \(B\), (c) range position \(C\), and (d) width (BW) plotted versus month.  Each year is shown as a line in a different color as indicated in the top-right corner of panel (a).

\begin{figure}
\includegraphics[width=11.0 cm,angle=90]{p1f2.pdf}
\caption{The variations in the echo band (a) amplitude of the fitted cosine function \(A\), (b) time shift or time of day of the peak \(B\), (c) average position of the band or the vertical shift of the cosine function from zero \(C\), and (d) width (BW). Different lines are for different years with the legend given in Figure \ref{fig:year_line}a.}
\label{fig:year_line}
\end{figure}

All four parameters peak in austral winter and decrease in summer months. Solar cycle variation is evident in (a) the band position amplitude and (c) the average band position. In general, both parameters were larger in years closer to solar minimum (year 2010) and decreased moving towards solar maximum (2013).  In both cases, the variation existed primarily in the winter months (around June). In contrast, the seasonal variation was also evident in (b) the band time shift and (d) the band width, but no clear solar cycle trend was present. In fact, the band time shift was quite consistent between different years, except for winter months. This parameter was typically around 15 hrs. This is when a maximum of \(R\left(t\right)\) was observed in UT, i.e. when band position was located furthest away from the radar.

In the previous analysis, the focus was on band characteristics related to range and their variations. The next presentation examines the diurnal/seasonal/solar cycle pattern as a whole for both band range position and one other important characteristic, the echo occurrence within the band. For each month, the band position \(R\left(t\right)\) and the average occurrence between \(R\left(t\right)\pm \sigma\) were found similar to those presented in Figure \ref{fig:month_avg_occ}. A horizontal slice of each panel of Figure \ref{fig:year_color} represents the time series for one month. The band position \(R\left(t\right)\) (average echo occurrence) is shown in Figures \ref{fig:year_color}a--\ref{fig:year_color}d (Figures \ref{fig:year_color}e--\ref{fig:year_color}h) as a color contour plot representing the time series for all months of the year simultaneously; different rows refer to  different years. The left panels look much smoother than the right ones; this is simply because each line on the left is a fitted cosine \(R\left(t\right)\), while on the right it is the average ``raw'' occurrence between two fitted cosine lines \(R\left(t\right)\pm \sigma\). The white data gaps at the bottom of Figures \ref{fig:year_color}d and \ref{fig:year_color}h reflect the fact that no MCM data were collected in January 2010.

\begin{figure}
\includegraphics[width=12.0 cm,angle=90]{p1f3.pdf}
\caption{Contour plots of the (a)--(d) echo band range and (e)--(h) occurrence versus UT and season. Each row shows the data for a different year as given in the top-right corner.}
\label{fig:year_color}
\end{figure}

In Figures \ref{fig:year_color}a--d, the band position \(R\left(t\right)\) peaks in nighttime (second half of the day) and in winter. This peak (yellow-to-red colors) is seen in the same plot area as the low occurrence plot cells in Figure \ref{fig:year_color}e--h (blue color). The amplitude of the range peak became consistently lower between 2010--2013.  This indicates that the high occurrence band moved closer to the radar for years further away from the solar minimum.

The minima in the band position \(R\left(t\right)\) (blue contours) were observed during the daytime and near equinoxes. This feature was most clear in 2010 when these minima were observed near 03 UT in March and September; these were also the two months presented in Figure \ref{fig:month_avg_occ}. The echo occurrence pattern, Figures \ref{fig:year_color}e--h, is less clear in terms of its maxima and their relationship with range minima. Typically, one equinoctial maximum was more pronounced as in 2011--2013, with no clear preference towards either March of September. In 2010, the two maxima are most similar (yellow contours), but the UT peak was shifted away from that of the range minimum (by about 5 hours). Overall, however, the range and occurrence patterns appear to be inverse of each other: larger range means smaller occurrence and vice versa.

The diurnal, seasonal, and solar-cycle variations in echo occurrence observed in Figures \ref{fig:month_avg_occ}--\ref{fig:year_color} can be explained by looking at the solar zenith angle (SZA) and ionospheric density conditions. SZA indicates how sunlit the ionosphere is and the associated degree of ionization, i.e. the \(E\)- and \(F\)-region electron densities, as explained in Section \ref{Sec:Intro}. The model NmE and NmF2 values have been employed for this analysis, Section \ref{Sec:Exp}. To keep model run times at a manageable level, here we consider 10-min cadence data including the average echo occurrence within the band (in Figures \ref{fig:month_avg_occ}--\ref{fig:year_color}, the 2-min data were considered).

Figures \ref{fig:scatter}a and \ref{fig:scatter}b present the results of this analysis for year 2010 in the form of scatter plots of band position versus (a) SZA and (b) model NmF2, while Figures \ref{fig:scatter}c and \ref{fig:scatter}d are scatter plots of echo occurrence versus (c) SZA and (d) model NmF2. In all four panels, the model NmE is shown by the color of the points, with the color bar given to the right. The points in the top row show some clustering in the form of closed loops (some of them appear just as horizontal lines). This is simply a consequence of the periodic nature of the cosine fit; each loop refers to the data for one month. The summer months exhibit little-to-no variation; these appear as horizontal loops-lines.

>>>>

% Chapter 2: paper1

@@

@@

\section{Introduction}
\label{Sec:Intro}

@@

Observations of decameter-scale irregularities with SuperDARN have an advantage of significant datasets accumulated under a wide variety of observational conditions. SuperDARN utilizes radio refraction in the High-Frequency (HF) band to observe \(F\) region echoes or backscatter from decameter-scale irregularities when radar beam achieves orthogonality with local magnetic field (orthogonality condition). Because of importance of propagation conditions for HF radar observations, the echo occurrence rates observed by SuperDARN are taken as a lower limit of the irregularity occurrence rates \citep{Chisham2007}.

Several recent SuperDARN studies have found clear seasonal and diurnal trends in the \(F\) region echo occurrence that formed a distinct pattern which appeared to be qualitatively similar to that of the solar zenith angle (SZA) variations \citep{Kane2012,Ghezelbash2014a,Ghezelbash2014b}. However, it has been challenging to determine how occurrence explicitly depends on SZA for all seasonal and time sectors. The nighttime occurrence, taken separately, was demonstrated to be closely related to SZA-controlled \(F\) region density  variations \citep{Kane2012}, but no similarly clear relationship has been demonstrated for datasets including daytime points.  One possible reason is the competing effects of the electric filed, particularly in the noon sector near the cusp region \citep{Ghezelbash2014a,Ghezelbash2014b}.

One difficulty is the need to consider propagation effects of plasma density. These arise because of \(F\)-region density controlling the amount of refraction and either underrefraction or overrefraction reducing echo occurrence when orthogonality condition is not met \citep{Milan1997,Danskin2002}. This is also related to how echo occurrence is generally considered a lower limit on irregularities, as described above. There is some ambiguity about whether changing irregularity production or detection by the radar is most responsible for any observed trends.

The production/detection ambiguity issue can be addressed by considering how both occurrence and distance from the radar (slant range) change in responses to various factors \citep{Kane2012,Ghezelbash2014b}. This is because range variations are expected to be almost entirely due to propagation effects and the differences in responses are therefore mostly likely due to changing production. In order to claim production as the dominant source of occurrence variability, occurrence analysis must explicitly consider echo location. In the current study this approach is employed to develop a comprehensive model of occurrence location, quantify echo location trends, and explicitly consider these trends when developing occurrence estimates. This is in contrast with previous studies that considered echoes detected in the entire (or fixed part) of the radar's field-of-view (FoV) \citep[e.g.][]{Kumar2011,Kane2012}.

@@

The primary instrument used in this study is the SuperDARN McMurdo radar located at McMurdo Station, Antarctica (77.88\(\deg\) S, 166.73\(\deg\) E, geographic) \citep[Figure 1]{Bristow2011}. Using a standard SuperDARN 3-letter radar code, this radar is referred to as MCM. SuperDARN is an array of coherent HF radars used to study processes in the Earth's ionosphere and magnetosphere. Radars are located at middle and high latitudes in both the Northern and Southern Hemispheres. Radars measure a 17-lag auto-correlation function (ACF) from which estimates of the Doppler velocity, power, and spectral width of ionospheric echoes in 75--100 range bins for each of the 16--24 radar beams are obtained. Each radar beam spans \(\sim3.25\deg\) horizontally and \(\sim30\deg\) vertically, with the entire FoV of \(\sim52\deg\)--\(78\deg\) in azimuth. Range bins typically start at 180 km and have a length of 15 km or 45 km and a radar scans through all beams in 60 s or 120 s depending on the mode of operation. Other details on operations and technical characteristics of SuperDARN radars are given by \citet{Chisham2007}.

The magnetic latitude (MLAT) of MCM is 80.0\(\deg\) S, based on the altitude-adjusted corrected geomagnetic (AACGM) coordinate system with 2010 coefficients (see recent paper by \citet{Shepherd2014}). The MCM FoV covers the southern polar cap and beam 14 of MCM is most conjugate with beam 1 of the SuperDARN Rankin Inlet (RKN) radar that overlooks the northern polar cap. The current study focuses on MCM observations, but in order to facilitate future comparisons with conjugate observations, the data from beam 14 of MCM are considered. MCM has been in routine operations since February 2010 and the current study has considered the data from the first 4 years of operations 2010--2013.

Standard SuperDARN criteria were used to eliminate ground scatter (\(\left| V-\Delta V \right| <30\) m/s and \(\left| W-\Delta W \right| <35\) m/s), where \(V\) is the Doppler velocity, \(W\) is the spectral width, and \(\Delta V\) and \(\Delta W\) represent uncertainties in these quantities. In addition, measurements with signal-to-noise ratio \(<\) 3 dB or \(W>\) 500 m/s were eliminated to reduce contribution from interference and other spurious echoes. In addition to standard operation modes, each SuperDARN radar has a set time periods where the radar is run in special modes. In order to have as continuous of a dataset as possible for the statistical analysis, data were used both from the standard modes and from special modes when possible. Modes where the beam 14 was over-sampled (beam camping modes) or not sampled at all were not used.

In this study, we also employed the IMF measurements taken from the high-resolution OMNI database using spacecraft-specific dataset for the WIND satellite. This was the OMNI dataset that had the largest number of valid measurements throughout the period of interest 2010--2013. All data in the OMNI database are time shifted from the spacecraft location to Earth's bow shock nose (BSN). The data were further shifted by 9--12 min to account for propagation from the BSN to the high-latitude ionosphere using a method described by \citet{Khan1999} that was also employed by \citet{Makarevich2012}. The IMF By and Bz components were averaged over 10-min intervals. Any periods with IMF information missing were excluded from the analysis.  In addition, the Polar Cap South (PCS) index data derived from magnetometer measurements at Vostock Antarctic station have been used \citep{Troshichev1988}.  The PC index is a standard index used to quantify responses of various polar cap phenomena to geomagnetic disturbance \citep[e.g.][]{Fiori2009}.

The model electron densities were also used in the current study. The electron densities were found from the International Reference Ionosphere (IRI) 2012 model \citep{Bilitza2014}. The IRI model is an empirical model of the ionosphere based on time of day, date, geographic location, altitude, and solar conditions, with the \(E\) region and \(F\) region peak electron densities (\(N_m E\) and \(N_m F_2\)) being of particular interest for this study.  The IRI model uses the International Union of Radio Science (URSI) or Consultative Committee on International Radio (CCIR) coefficient maps of the critical frequencies ot produce \(N_m E\) and \(N_m F_2\), with both the URSI and CCIR options implemented in the current study.  In addition, IRI 2012 has three options for the bottomside thickness parameter B0, with all three options used in the current study.

@@

In order to examine diurnal, seasonal, and solar cycle trends in echo occurrence, the monthly average occurrence was found for each universal time (UT) 2-min interval and for each 45-km range bin. Figures \ref{fig:month_avg_occ}a--\ref{fig:month_avg_occ}d show monthly average occurrence expressed in percentage for four selected months in year 2010 versus UT and slant range. The four months presented were chosen to be representative of different seasons. Figures \ref{fig:month_avg_occ}e--\ref{fig:month_avg_occ}h on the right similarly shows the monthly average peak \(F\)-region electron density \(N_mF_2\) obtained from the IRI model using its default option of the URSI coefficient maps and the bottomside thickness B0 from \citet{Altadill2009}.  For the other five IRI options, the plasma density profiles were also calculated and used in ray tracing calculations, as discussed below.  The \(N_mF_2\) values for these other five options are not presented here.  As compared with the IRI 2012 default option, the average percent difference for the three CCIR options were 15\%--22\% in March, June, and September and \(\sim9\)\% in December.  For the three URSI options, they were 3\%--9\% in March, June, and September and \(\sim2\)\% in December.  Finally, the cosine white contours in both Figures \ref{fig:month_avg_occ}a--\ref{fig:month_avg_occ}d and \ref{fig:month_avg_occ}e--\ref{fig:month_avg_occ}h show the approximate range position of the \(F\)-region echo band for each month determined as described below.

\begin{figure}
\includegraphics[width=12.5 cm,angle=90]{p1f1.pdf}
\caption{Monthly average plots of (a--d) echo occurrence and (e--h) peak electron density for 2010.  Four months are selected (one for each season) to show seasonal variations. In Figures \ref{fig:month_avg_occ}a--\ref{fig:month_avg_occ}d,  the pink line shows the 25\% occurrence contour that is used to define the echo band region. Black dots are the midpoints between the upper and lower portion of this contour.  A cosine function (white line) is fit to these points and used to model the echo band. The dashed white lines are the upper and lower bounds of the model band. The top right corner of each panel shows the parameters that characterize the band for that month: \(A\) is amplitude of cosine function, \(B\) is time shift, \(C\) is range shift from 0, and BW is the band width.  In Figures \ref{fig:month_avg_occ}e--\ref{fig:month_avg_occ}h the same echo location model is also given (white lines) along with expected echo locations based on ray tracing simulations (solid lines).  The red (grey) lines refer to three URSI (CCIR) options (see text for details).}
\label{fig:month_avg_occ}
\end{figure}

Figures \ref{fig:month_avg_occ}a--\ref{fig:month_avg_occ}d show a distinct \(F\)-region band of high occurrence in every month at ranges 500--1500 km. Some \(E\)-region echoes are also observed at shorter ranges 180--300 km in June at 01--06 UT and in December at 17--24 UT.  In addition, Figures \ref{fig:month_avg_occ}a, \ref{fig:month_avg_occ}c, and \ref{fig:month_avg_occ}d show some echoes with low-occurrence near 04 UT.  These observations refer to the region just equatorward of the polar cap near the noon sector (looking through the polar cap) and thus may represent a radar signature of the dayside cusp \citep{Yeoman2002}.  In the following analysis, we focus on the maid \(F\)-region band between 500 km and 1500 km which refers to observations in the polar cap.

The \(F\)-region band range position and peak occurrence within the band was changing most clearly throughout the day in March and September, i.e. diurnal variation in both parameters is observed clearly in spring and fall months. To highlight the high echo occurrence band, a contour of 25\% occurrence is plotted in each Figure \ref{fig:month_avg_occ}a--\ref{fig:month_avg_occ}d by the pink line.  The midpoints were also found between the top and bottom parts of the pink 25\% occurrence contour.  These are plotted by the small black dots and referred to below as the band position \(r(t)\), where range \(r\) is in kilometers and time \(t\) is in hours.

Both the 25\% occurrence contours and the band position \(r(t)\) exhibited a clear cosine-like variation with a 24-h period in March and September 2010.  Very similar variations were observed in other years and other near-equinoctial months.  Further, similar variations were also observed for many individual days.  IN summer months, the variation was largely flat, Figure \ref{fig:month_avg_occ}d, which can be also represented by a cosine variation shifted from zero with a small amplitude.  To further characterize this diurnal variation, a cosine function of the form
\begin{equation}\label{E:Model}
R\left(t\right) = A \cos\left[2\pi\left(t-B\right)/24\right] + C
\end{equation}
was fitted to the band position values \(r(t)\). Here \(R(t)\) is the fitted range of the center of the band in kilometers, and  \(A\), \(B\), and \(C\) are fit coefficients with units of kilometers, hours, and kilometers, respectively. The white solid line in Figures \ref{fig:month_avg_occ}a--\ref{fig:month_avg_occ}h shows the resulting fit \(R\left(t\right)\). The white dashed lines represent the upper and lower limits of the band.  They were calculated by first finding the average distance \(\sigma\) between points in the 25\% occurrence contour and the fitted cosine at the same time of day. The original cosine fit was then shifted up or down by this amount as \(R\left(t\right) \pm \sigma\) to give the upper and lower limits of the band.  The bandwidth (BW) could then be defined as \(2\sigma\).  The fit parameters and band width are given in the top right corner of Figures \ref{fig:month_avg_occ}a--\ref{fig:month_avg_occ}d. For the rest of this study, ``echoes within the band'' refer to data points that fall between the upper and lower white dashed lines. From Figures \ref{fig:month_avg_occ}a--\ref{fig:month_avg_occ}d, both the center position and upper/lower bounds for the band shown by the white lines capture the band and its diurnal variation in range very well for all months. Therefore, these estimates are used below to quantify the echo band behavior, particularly in context of expected electron densities.

MCM is oriented in such a way that it samples daytime (nighttime) ionosphere in roughly the first (second) half of the day in UT. The electron density profiles change significantly between the nighttime and daytime periods, with one important consequence being higher peak densities \(N_mF_2\) during the day.  This is illustrated in Figure \ref{fig:month_avg_occ}e--\ref{fig:month_avg_occ}h that clearly shows higher daytime \(N_mF_2\) values that reach a peak near 05 UT at farther ranges and lower nighttime \(N_mF_2\) values that reach a minimum near 17 UT.  The other important effect is an increase in the peak height \(h_mF_2\) during the night (not presented here).

The \(F\)-region echo band was also closer to the radar in the daytime and farther away at night. This feature is most clear for the March and September data sets. It is much less clear in June, although one can still see a portion of the cosine-like variation at 02--10 UT in June in the pink contour. In this month, the white contours are mostly extrapolation outside of this interval, but they fit into the overall diurnal/seasonal pattern well (see also next analysis).  In December, the band is flat and diurnal variation is not evident.

In terms of diurnal variation of echo occurrence, there was a minimum occurrence at nighttime and a maximum in daytime hours. The minimum was much deeper and more extended in June so that essentially no echoes were observed at 12--24 UT. At this time, the expected location of echoes based on the cosine-fit model (white lines) is farther away from MCM than in any month and any UT interval within uncertainty, but no echoes are observed either at these ranges or any other ranges. This is in contrast with March and September, where occurrence is reduced at nighttime but not to zero. Overall, Figure \ref{fig:month_avg_occ} shows that range position of the band is largely controlled by the density conditions, while the echo occurrence variation is perhaps more complicated.

The final observation about Figure \ref{fig:month_avg_occ} concerns the results of raytracing simulations that were conducted for each hour of day 15 in each of the four months shown in Figure \ref{fig:month_avg_occ} using a version of standard ray tracing tools based on numerical solutions of the Hamiltonian raypath equations \citep{Haselgrove1963,Jones1975} that allow the consideration of arbitrary density distributions. In these simulations, IRI densities in a 2-D simulation domain aligned with beam 14 of MCM were used.  That is, plasma densities employed were varying both horizontally and vertically.  The grid bins in the horizontal and vertical directions were 5 km and 1 km, respectively, which is well below the radar cell size of 45 km.  Rays separated by \(1\deg\) in elevation were found under the assumption of ordinary mode propagation.  Locations were founds where the ray was nearly perpendicular (within \(1\deg\)) to the local magnetic field based on the International Geomagnetic Reference Field 12 model \citep{Thebault2015}.

The range locations where orthogonality condition was expected are plotted in Figures \ref{fig:month_avg_occ}e--\ref{fig:month_avg_occ}h by six lines corresponding to six different options of the IRI model described above.  The red (grey) kines refer to three URSI (CCIR) options.  These lines are seen between the solid and bottom dashed white lines, indicating a reasonable agreement between the expected and observed backscatter locations.  The important exception is seen, however, for nighttime periods (low-density periods).  During these periods, no orthogonality was expected based on IRI densities, but significant (albeit reduced) amount of backscatter was observed in March and September.  This implies a significant source of nighttime plasma density that is unaccounted for by the IRI model.  IRI is known to underestimate \(N_mF_2\) during equinoctial nighttime periods in the northern polar cap \citep{Themens2014}, and our observations suggest that a similar underestimation also occurs in the southern polar cap.  The UT extent of the period with no expected orthogonality is reduced in the CCIR option, which suggests a better ionospheric representation in this option.  This result is also consistent with that reported by \citet{Themens2014}.

The same analysis was performed for the entire period of interest 2010--2013 and the band characteristics are presented in Figure \ref{fig:year_line}. It shows the band position amplitude \(A\), Figure \ref{fig:year_line}a, time shift \(B\), Figure \ref{fig:year_line}b, range position \(C\), Figure \ref{fig:year_line}c, and width (BW), Figure \ref{fig:year_line}d, plotted versus month.  Each year is shown as a line in a different color as indicated in the top-right corner of Figure \ref{fig:year_line}a.  All four parameters peak in austral winter and decrease in summer months. Solar cycle variation is evident in the band position amplitude, Figure \ref{fig:year_line}a, and the average band position, Figure \ref{fig:year_line}c. In general, both parameters were larger in years closer to solar minimum (year 2010) and decreased moving towards solar maximum (2013).  In both cases, the variation existed primarily in the winter months (around June). In contrast, the seasonal variation was also evident in the band time shift, Figure \ref{fig:year_line}b, and the band width, Figure \ref{fig:year_line}d, but no clear solar cycle trend was present. In fact, the band time shift was quite consistent between different years, except for winter months. This parameter was typically around 15 hrs. This is when a maximum of \(R\left(t\right)\) was observed in UT, i.e. when band position was located furthest away from the radar.

\begin{figure}
\includegraphics[width=11.0 cm,angle=90]{p1f2.pdf}
\caption{The variations in the echo band (a) amplitude of the fitted cosine function \(A\), (b) time shift or time of day of the peak \(B\), (c) average position of the band or the vertical shift of the cosine function from zero \(C\), and (d) width (BW). Different lines are for different years with the legend given in Figure \ref{fig:year_line}a.  The fit errors are given by the vertical bars.}
\label{fig:year_line}
\end{figure}

In the previous analysis, the focus was on band characteristics related to range and their variations. The next presentation examines the diurnal/seasonal/solar cycle pattern as a whole for both band range position and one other important characteristic, the echo occurrence within the band. For each month, the band position \(R\left(t\right)\) and the average occurrence between \(R\left(t\right)\pm \sigma\) were found similar to those presented in Figure \ref{fig:month_avg_occ}. A horizontal slice of each panel of Figure \ref{fig:year_color} represents the time series for one month. The band position \(r\left(t\right)\) is shown in Figures \ref{fig:year_color}a--\ref{fig:year_color}d as a color contour plot representing the time series for all months of the year simultaneously; different rows refer to  different years. Similarly, the average echo occurrence is shown in Figures \ref{fig:year_color}e--\ref{fig:year_color}h.  Figures \ref{fig:year_color}a--\ref{fig:year_color}d also present the fitted band position R(t) as line contours using the same colour scheme.

The white gaps appear in Figures \ref{fig:year_color}a--\ref{fig:year_color}d during the periods with no estimates for \(r(t)\).  For the fitted band position \(R(t)\), these gaps are filled except for January 2010 when no MCM data were available.  Outside the gaps, the \(R(t)\) contours match well the underlying color pattern.  INside the gaps, they represent extrapolation of the same pattern. This information is alos useful ot analyze, particularly considering that the fitted errors for the months that contibute the most to the gaps were less tan 20\%, Figures \ref{fig:month_avg_occ}b and \ref{fig:year_line}.
